Procurements of the Canadian Government, Runaway Military Contracts

### 2024-12-03 paper/paper.qmd
new color scheme

#colors!
def_c <- "#e41a1c"
fish_c <- "#377eb8"
health_c <- "#4daf4a"
services_c <- "#ff7f00"
public_c <- "#984ea3"
space_c <- "#ffff33"

@@

# Create the grouped bar chart
ggplot(data = grouped_df, aes(x = year, y = total_spending_billion, fill = buyer_category)) +
  geom_bar(
    stat = "identity",
    position = position_dodge(width = 0.9), # Adjust dodge width
    width = 0.8 # Adjust bar width
  ) +
  scale_fill_manual(
    values = c(
  "National Defence" = def_c,
  "Public Services and Procurement Canada" = public_c,
  "Shared Services Canada" = services_c,
  "Public Health Agency of Canada" = health_c,
  "Fisheries and Oceans Canada" = fish_c,
  "Other"  = "gray"
    )
  ) +
  scale_y_continuous(
    labels = comma_format(accuracy = 1), # Formats y-axis labels without decimal places
    breaks = pretty_breaks(n = 5), # Ensures approximately 5 breaks like 5, 10, 15
    expand = expansion(mult = c(0, 0.05)) # Adds a small space above the highest bar
  ) +
  labs(
    x = "Year",
    y = "Total Spending (Billions)",
    fill = "Buyer",
    caption = "."
  ) +
  theme_minimal() +
  theme(
    legend.text = element_text(size = 8), # Adjusts the size of legend text
    legend.title = element_text(size = 9), # Adjusts the size of legend title
    legend.key.size = unit(0.6, "lines") # Reduces the size of legend symbols
  )

@@

```{r fig-top-supps}
#| message: false
#| echo: false
#| warning: false
#| fig-cap: Top 10 Suppliers by Total Amount Awarded
#| fig-align: center
#| fig-pos: "H"
# Merge total aggregated amounts back into aggregated_df
aggregated_df <- aggregated_df %>%
  left_join(aggregated_totals, by = "word", suffix = c("", "_total"))

@@

# Use this ordering in the plot
# Use this ordering in the plot (reversed order)
ggplot(aggregated_df_top10, aes(x = reorder(word, total_amount_total), y = total_amount / 1e9, fill = buyer)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(
    x = "Supplier",
    y = "Total Amount Awarded (in Billions)",
    fill = "Buyer"  # Adds a title to the legend
  ) +
  scale_y_continuous(labels = scales::comma) +  # Format y-axis with commas
  scale_fill_manual(
    values = c(
      "National Defence" = def_c, 
      "Canadian Space Agency" = space_c, 
      "Shared Services Canada" = services_c, 
      "Public Health Agency of Canada" = health_c, 
      "Fisheries and Oceans Canada" = fish_c,
      "Other" = "gray"  # Assign a default color for other buyers if necessary
        ), 
    breaks = selected_buyers  # Restricts legend to selected buyers
  ) +
  theme_minimal() +
  theme(
    legend.position = "right"  # Positions the legend on the right
  )

>>>>

#colors!
def_c <- "#fb8072"
fish_c <- "#80b1d3"
health_c <- "#4daf4a"
services_c <- "#fdb462"
public_c <- "#bebada"
space_c <- "#ffffb3"

@@

# Create the grouped bar chart
ggplot(data = grouped_df, aes(x = year, y = total_spending_billion, fill = buyer_category)) +
  geom_bar(
    stat = "identity",
    position = position_dodge(width = 0.9), # Adjust dodge width
    width = 0.8 # Adjust bar width
  ) +
  scale_fill_manual(
    values = c(
  "National Defence" = def_c,
  "Public Services and Procurement Canada" = public_c,
  "Shared Services Canada" = services_c,
  "Public Health Agency of Canada" = health_c,
  "Fisheries and Oceans Canada" = fish_c,
  "Other"  = "gray"
    ),
  guide = guide_legend(nrow = 3)
  ) +
  scale_y_continuous(
    labels = comma_format(accuracy = 1), # Formats y-axis labels without decimal places
    breaks = pretty_breaks(n = 5), # Ensures approximately 5 breaks like 5, 10, 15
    expand = expansion(mult = c(0, 0.05)) # Adds a small space above the highest bar
  ) +
  labs(
    x = "Year",
    y = "Total Spending (Billions)",
    fill = "Buyer",
    caption = "."
  ) +
  theme_minimal() +
  theme(legend.position = "bottom"
  )

@@

```{r fig-top-supps}
#| message: false
#| echo: false
#| warning: false
#| fig-cap: Top 10 Suppliers by Total Amount Awarded
#| fig-align: center
#| fig-pos: "H"

# Create a custom mapping for supplier renaming
supplier_rename <- c(
  "cae" = "CAE", 
  "skyalyne" = "Skyalyne", 
  "vancouver shipyards" = "Vancouver Shipyards", 
  "f35" = "F-35 Joint Program", 
  "bell" = "Bell", 
  "airbus" = "Airbus", 
  "textron" = "Textron Avaition", 
  "dynamics" = "General Dynamics",
  "dettwiler" = "Dettwiler and Associates", 
  "atomics" = "Atomics Aeronautical Systems"
)

selected_buyers <- c(
  "National Defence", 
  "Canadian Space Agency", 
  "Shared Services Canada", 
  "Public Health Agency of Canada", 
  "Fisheries and Oceans Canada"
)

# Adjust the dataframe with supplier renaming
aggregated_df <- aggregated_df %>%
  mutate(word = ifelse(word %in% names(supplier_rename), supplier_rename[word], word))

aggregated_totals <- aggregated_df %>%
  group_by(word) %>%
  summarise(total_amount = sum(total_amount, na.rm = TRUE)) %>%
  arrange(desc(total_amount))

# Merge total aggregated amounts back into aggregated_df
aggregated_df <- aggregated_df %>%
  left_join(aggregated_totals, by = "word", suffix = c("", "_total"))

@@

# Use this ordering in the plot
# Use this ordering in the plot (reversed order)
ggplot(aggregated_df_top10, aes(x = reorder(word, total_amount_total), y = total_amount / 1e9, fill = buyer)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(
    x = "Supplier",
    y = "Total Amount Awarded (in Billions)",
    fill = "Buyer"
  ) +
  scale_y_continuous(labels = scales::comma) +  # Format y-axis with commas
  scale_fill_manual(
    values = c(
      "National Defence" = def_c, 
      "Canadian Space Agency" = space_c, 
      "Shared Services Canada" = services_c, 
      "Public Health Agency of Canada" = health_c, 
      "Fisheries and Oceans Canada" = fish_c,
      "Other" = "gray"  # Assign a default color for other buyers if necessary
        ), 
    breaks = selected_buyers,
    guide = guide_legend(nrow = 2)
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom"  # Positions the legend on the right
  )
